---
title: "ICU + Severity Example for `wsre`"
author: "Andrew Manderson"
date: "`r format(Sys.time(), '%d %B, %Y')`"
fontfamily: tgpagella
fontsize: 10pt
papersize: a4
geometry: margin=2.25cm
bibliography: ../0bibliography/year-1-bib.bib
csl: aam71-test.csl
output: 
  pdf_document:
    includes:
      in_header:
        tex-input/pre.tex
    fig_caption: true
    number_sections: true
    keep_tex: true
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, comment = NA, out.width = "85%", fig.align = "center", auto_pdf = TRUE)
```

# Model origins

@presanis:etal:14 undertook a large evidence synthesis in order to estimate the severity of the H1N1 pandemic amongst the population of England.
The H1N1 pandemic occurred in three _waves_, and the change in severity between waves is of interest.
Due to the time between waves, the available sources of data differ between waves.
During the third wave some data sources underwent a decrease in typical sample size, and began aggregating previously available data in a coarser manner.
To ameliorate the impact of the data coarsening, independent data on one aspect of the large severity model, the number of suspected influenza cases in hospital's intensive care units (ICU), can be incorporated.
@goudie:etal:18 presents a Markov melding approach for joining the independent ICU model with a simplified version of the larger severity model.
We consider this same pair of models here, as the melded parameter has an unknown prior marginal distribution under both models, and the independent nature of the data sources could result in a degree of conflict between the various intermediary melding distributions.
This is also an example of the Markov melding where the melded model has no obvious implied joint model, as the link parameter is not a root node, in a DAG representation, for either model.
The code to reproduce all figures and outputs from this experiment is available at https://github.com/hhau/full-melding-example.

## ICU

The data for the ICU submodel ($\modelindex = 1$) considers aggregate weekly counts of patients in the ICU of all the hospitals in England.
Such data is known to be an underestimate of the true number of patients affected by H1N1 influenza, and also requires careful temporal consideration, as patients enter and exit the ICU in a stochastic manner.
The third wave of the H1N1 influenza predominantly occurred during a 78 day block between December 2010 and February 2011.
ICU observations were recorded weekly on days $U = \{8, 15, \ldots, 78\}$, and considered the number of children $a = 1$ and adults $a = 2$ in the ICU on the day of observation.
Denote a specific weekly observation as $y_{a, t}$ for $t \in U$.
The estimated lower bound on the cumulative number of ICU admissions over the whole 78 day block $T = \{1, 2, \ldots, 78\}$, for each age group, functions as our link parameter $\phi = (\phi_{1}, \phi_2)$.
A need to appropriately model the temporal nature of the weekly ICU data motivates a time inhomogeneous Poisson process.
The process has time inhomogeneous rate parameter $\lambda_{a, t}$ for $t \in T$, with an age group specific ICU exit rate $\mu_{a}$, which is exponentially distributed.

To address the difference between the observation times $U$, and the daily support of our Poisson process $T$, the expected number of influenza patients in the ICU at time $t$ is
\input{tex-input/icu-model/0010-thinned-poisson-process.tex}
An identifiability assumption of $\eta_{a, 1} = 0$ is required, which reasonably assumes that no H1N1 influenza patients were in the ICU at time $t = 0$.

Influenza strains do not exist in isolation, hence not all ICU admissions with influenza symptoms were the result of the H1N1 pandemic.
Consider $\lambda_{a, t}$ as the expected number of _suspected_ H1N1 cases.
Weekly virological positivity data is available at weeks $V = \{1, \ldots, 11\}$, and informs the proportion of influenza cases which are attributable to the H1N1 virus $\pi_{a, t}^{\text{pos}}$.
The virology data consists of the number of H1N1-positive swabs $z_{a, v}^{\text{pos}}$ and the total number of swabs tested for influenza that week $n_{a, z}^{\text{pos}}$.
This proportion relates the counts to $\pi_{a, t}^{\text{pos}}$ a truncated uniform prior on $\pi_{a, t}^{\text{pos}}$,
\input{tex-input/icu-model/0020-pipos-unif-model.tex}
with $v = 1$ for $t = 1, 2, \ldots, 14$, and $v = \lfloor(t - 1) \mathop{/} 7 \rfloor$ to align the temporal indices.
The estimate of $\pi_{a, t}^{\text{pos}}$ is then used to compute $\phi_{a}$
\input{tex-input/icu-model/0021-phi-icu-calc.tex}
This summation is a non-invertible function, which guides our choice of model ordering when melding, which we will discuss in Section&nbsp;\ref{discussion}. 

Lastly, we specify priors for the remaining parameters. 
A lognormal random walk is used for the expected number of new admissions 
\input{tex-input/icu-model/0030-prior-set-one.tex}
for $t = 2, 3, \ldots, 78$ and $a = {1, 2}$.
Age group specific exit rates have informative priors specified for them in an identifiable manner 
\input{tex-input/icu-model/0040-prior-set-two.tex}
The lower bound on the positivity proportion has a flat prior
\input{tex-input/icu-model/0041-prior-positivity-proportion.tex}

## Severity

A simplified version of the remaining components of the large severity model of @presanis:etal:14 ($\modelindex = 2$) is considered here.
The estimate of the cumulative number of ICU admissions, $\phi_{a}$, is assumed to be an underestimate of the true number of of ICU admissions due to H1N1, $\chi_{a}$.
The model we consider to address this is 
\input{tex-input/icu-model/0050-sev-submodel.tex}
where $\pi^{\text{det}}$ is the age group constant probability of detection, and the priors on $\chi_{a}$ are appropriate summaries of the other components of the larger severity model.

# Melding

## Prior, stage one comparison

```{r prior_comparison, fig.cap = "Left to right: Severity prior $\\pd_{2}(\\phi)$, ICU prior $\\pd_{1}(\\phi)$, and the stage one (ICU) posterior  $\\pd_{1}(\\phi \\mid Y_{1})$."}
knitr::include_graphics("plots/prior-stage-one-comparison.pdf")
```

### Priors

In both the severity and ICU submodels, the prior marginal distribution $\pd_{\modelindex}(\phi)$ is unknown.
As noted in @goudie:etal:18, and visible in the centre panel of Figure&nbsp;\ref{fig:prior_comparison}, the prior marginal distribution for the ICU submodel, $\pd_{1}(\phi)$, is effectively flat in the region of interest (ROI).
This implies that the self-density ratio of two points in the ROI is indistinguishable from 1.
Applying our weighted-sample self-density ratio estimation method is unlikely to result in an improvement in accuracy for $\hat{\pd}_{1}(\phinu) \mathop{/} \hat{\pd}_{1}(\phide)$.
Conversely, the severity model has substantially more curvature in the ROI, and our methodology is likely to improve the accuracy of the ratio estimate where one or both of the points is far away from the prior mean.

### Stage one

In stage one we target $\pd_{1}(\phi, \psi_{1} \mid Y_{1}) \propto \pd_{1}(\phi, \psi_{1}, Y_{1})$.
These samples for $\phi$ are displayed in the right panel of Figure&nbsp;\ref{fig:prior_comparison}, and we see that whilst there is substantial overlap with $\pd_{2}(\phi)$ (left panel), $\pd_{1}(\phi, \mid Y_{1})$ is more disperse, particularly for $\phi_{1}$.
The choice to target the ICU posterior enables the use of the original `JAGS` [@plummer:18] implementation of @goudie:etal:18 in order to sample the posterior of interest.
Note that this is a different stage one target than the HIV example, and demonstrates that we are free to choose which components of the melded posterior comprise the stage one target.
<!-- We are also opting to sample the ICU posterior first, as the link function in Equation&nbsp;\ref{eqn:phi-icu-calc} is non-invertible, and appropriately extending it to an invertible function is challenging.
 -->

## Estimating the severity self-density ratio for stage two

### Practicalities

As we have targeted $\pd_{1}(\phi, \psi_{1} \mid Y_{1})$ in stage one, the stage two acceptance probability for a Gibbs update for $\phi$ is
\input{tex-input/icu-model/0060-icu-stage-two-acceptance-probability.tex}
Our focus in stage two is on the accurate estimation of $\pdrh_{2}(\phinu, \phide) = \pd_{2}(\phinu) \mathop{/} \pd_{2}(\phide)$.
We use a grid of 100 points for the mean of the Gaussian weighting functions $\mu_{\wfindex}$ based on taking 10 equally spaced values, between 30 and 275 for $\phi_{1}$, and 500 and 3000 for $\phi_{2}$, and $\sigma^2_{\wfindex} = (25^2, 250^2)$ for all $\wfindex$.
Each weighted target has 1000 MCMC samples drawn from it, for 100000 MCMC samples in total. 
All individual ratio estimates are used to build an improved estimate for $\pdrh_{2}(\phinu, \phide)$ using the method described in Section&nbsp;\ref{sec:multiple-weighting-functions}.
For a fair comparison, we draw 100000 samples from $\pd_{2}(\phi)$ which we use to estimate $\hat{\pd}_{2}(\phi)$ for use in Equation&nbsp;\eqref{eqn:icu-stage-two-acceptance-probability}, and compare the melded posteriors produced using these two estimates.

### Stage two trace plots

```{r stage_two_phi_traces, fig.cap = "Stage two trace plots for $\\phi$ using the KDE estimate in $\\alpha(\\phi^{*}, \\phi)$ (left column) and the weighted-sample self-density ratio estimate (right column)."}
knitr::include_graphics("plots/stage-two-traces.pdf")
```

Figure&nbsp;\ref{fig:stage_two_phi_traces} displays trace plots of the MCMC chains targeting the stage two target, where $\alpha(\phi^{*}, \phi)$ is computed using the naive KDE (left column), and the weighted-sample self-density ratio estimate (right column).
From the erroneous behaviour displayed in the left column, we can infer that the KDE has produced an underestimate in the tails.
This leads to an acceptance probability that is much larger than it should be, as the proposal term is in the denominator of Equation&nbsp;\eqref{eqn:icu-stage-two-acceptance-probability}.
Hence, moves to improbable values of $\phi$ are associated with acceptance probabilities that are dominated by epistemic error in the KDE.
Once at this improbable value, the KDE error then has the opposite effect, as the underestimate now results in the chain being unable to move back to probable values.
The produces the monotonic step-like behaviour seen in the left column of Figure&nbsp;\ref{fig:stage_two_phi_traces}.
It would be reasonable to argue that given the visible behaviour in the bottom right panel of Figure&nbsp;\ref{fig:stage_two_phi_traces}, and usual suite of convergence criteria, the stage two sampler needs to be run for longer.
However, the fact that the poor visual properties of the trace plots, and the associated numerical convergence criteria, could be ameliorated with additional sampling is encouraging, as this is not the case for the "No WSRE" sampler.

# ICU example conclusion

Melding without employing the weighted-sample self-density ratio estimation (WSRE) idea results in clearly erroneous behaviour in the ICU example.
Whilst there is no baseline "truth" to reference in this example, as the implied joint distribution is opaque and not easily sampled, the sampler that employs the WSRE method for $\pd_{2}(\phi)$ produces feasible results, where as the naive KDE approach does not.
This improved behaviour is obtained using the same number of samples from the prior marginal distributions, or weighted versions thereof.

<!-- -------------------- END OF MAIN BODY OF DOCUMENT -------------------- -->
\newpage

<!-- The {-} tag here suppresses the section numbering. -->
# Bibliography {-}

<!-- This makes pandoc-citeproc put the references before the end of document. -->
<div id="refs"></div>

\newpage

<!-- Now switch to alphabetical numbering for the appendix, and reset the counter. -->
\renewcommand{\thesection}{\Alph{section}}
\setcounter{section}{0}

# Appendix 