---
title: "ICU + Severity Example for `wsre`"
author: "Andrew Manderson"
date: "`r format(Sys.time(), '%d %B, %Y')`"
fontfamily: tgpagella
fontsize: 10pt
papersize: a4
geometry: margin=2.25cm
bibliography: ../0bibliography/year-1-bib.bib
csl: aam71-test.csl
output: 
  pdf_document:
    includes:
      in_header:
        tex-input/pre.tex
    fig_caption: true
    number_sections: true
    keep_tex: true
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, comment = NA, out.width = "85%", fig.align = "center", auto_pdf = TRUE)
```

# Introduction

High level summary:

Two models arising from different data sources / scopes.
Not obvious how to join them into one big model.
Non-invertible link function makes the ICU model a clear choice for stage one.


## Origins of the models

### ICU

### Severity

## How best to join? 

Don't know if this is necessarily a paragraph/section.

## Model ordering

- Here, there is a natural ordering due to the non-invertible link function.
- In general it doesn't mathematically matter about this, because it can be extended into an invertible one, but in practice the performance of the multi-stage sampler heavily depends on the choice of link function extension.

- The natural ordering is also widest / least informative model first. 
- This way we are least sensitive to differences in location between the models.
- The cost is that we are susceptible to particle degeneracy / sample shrinkage.
  - This is a problem we can monitor though, differences in location we cannot. 

# Compare the Priors

See Figure&nbsp;\ref{fig:prior_comparison}.

```{r prior_comparison, fig.cap = "Prior comparison, should look like the paper."}
knitr::include_graphics("plots/prior-comparison.pdf")
```

# Sample the ICU Posterior

See Figure&nbsp;\ref{fig:icu_post_plot}.

```{r icu_post_plot, fig.cap = "ICU model posterior, same scales as prior."}
knitr::include_graphics("plots/icu-posterior.pdf")
```


# Estimate the severity self-density ratio using `wsre`

We use a grid of values for $\phi_{1}$ and $\phi_{2}$ based on taking 10 equally spaced values, between 30 and 275 for $\phi_{1}$, and 500 and 3000 for $\phi_{2}$, and taking the outer product of these values.

- Is is possible to visualise this estimate? Parallel coordinates? Four input dimensions and an output dimension is hard.

# Melding


## Pooled priors

- Eventually I would like to compare linear / log pooling.
    - It's an easy, drop in replacement.

## Compare the melded output with/without wsre

### With `wsre`

$\phi$ trace:

```{r phi_trace_wsre, fig.cap = "$\\phi$ trace"}
knitr::include_graphics("plots/stage-two-wsre-phi-trace.pdf")
```

$\psi_{2}$ trace:

```{r theta_tr}
knitr::include_graphics("plots/stage-two-wsre-theta-trace.pdf")
```

<!-- TODO: Change theta to phi
 -->

Something is blowing up and pining the $\psi_{2}$ traces to 1 or 0.
This is a numerical instability, I think its why Rob wrote the numerically stable continuous extension to the binomial? This must be an issue inside the `lgamma` function or the like

# Conclusion

Concise summary

<!-- -------------------- END OF MAIN BODY OF DOCUMENT -------------------- -->
\newpage

<!-- The {-} tag here suppresses the section numbering. -->
# Bibliography {-}

<!-- This makes pandoc-citeproc put the references before the end of document. -->
<div id="refs"></div>

\newpage

<!-- Now switch to alphabetical numbering for the appendix, and reset the counter. -->
\renewcommand{\thesection}{\Alph{section}}
\setcounter{section}{0}

# Appendix 